---
title: "Writing Reports with R Markdown"
author: "Kel S"
date: "2023-03-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
library(tidyverse)
```

This report was prepared to the attention of the United Nations. It analyzes the relationship between a country's GDP, its life expectancy, and CO2 emission. Our goal is to determine to what degree a country's economic strength or weakness may be related to its public health status and impact on climate pollution. We hypothesize that both life expectancy and CO2 emission increase with a country's GDP.


```{r data, include=FALSE}
gapminder_1997 <- read_csv("data/gapminder_1997.csv")
```

```{r gdp_lifeexp_1997, echo=FALSE}
ggplot(data = gapminder_1997) +
  aes(x = gdpPercap, y = lifeExp, color = continent, size = pop/1000000) +
  geom_point() +
  labs(x = "GDP Per Capita",
       y = "Life Expectancy",
       title = "Do people in wealthy countries live longer?",
       size = "Population (in millions)")

```

```{r variables, include=FALSE}
nCountries <- gapminder_1997 %>%
  select(country) %>%
  n_distinct()

minGDP <- gapminder_1997 %>%
  summarise(round(min(gdpPercap)))%>%
  pull()

maxGDP <- gapminder_1997 %>%
  summarise(round(max(gdpPercap)))%>%
  pull()

```


The above plot shows the relationship between GDP per capita and life expectancy for a total of `r nCountries` countries. This set of countries, economic wealth ranged from a minimum of USD `r minGDP` to a maximum of USD `r maxGDP` per capita. 



|Summary of Data|
|--------|------|
|Number of countries|`r nCountries`|
|Maximum GDP per capita|`r maxGDP`|
|Minimum GDP per capita|`r minGDP`|

```{r kabletable, echo=FALSE}
library(knitr)
gapminder_1997 %>%
  filter(continent == "Oceania") %>%
  kable()
  
```


# Header 1
## Header 2
### My third header

Exercise: R  Markdown Syntax
* Create a bullet point list with three items
* as the first item, write the name of your currenlty favorite programming language in bold
* as the second item, write the name of a function you have so far found most useful in italics
* as the third item write one thing you want to learn next on your programming journey in italics
* turn your bullet point list into a numbered list
* create a fourth list item and find an online guide and/or cheat sheet for a basic Markdown syntax, write its name down here and hyperlink its url

 My List:
- **R** 
- *select()*
- ***how to make a lab notebook in R***

1. **R**
1. *select()*
1. *how to make a lab notebook in R*
1. [**Markdown Cheatsheet**](https://www.markdownguide.org/cheat-sheet/) 


Steps for a git flow:

1. git status
1. git add
1. git commit -m "message"
1. git push

